Reasoning about a Capability Machine with Local Capabilities\\ Provably Safe Stack and Return Pointer Management (without OS Support)

### 2017-02-13 csf-2017/paper.tex
updates to the new figure
Rewrote some more about g1

In Lemma~\ref{lem:correctness-g1} the order of control has been
inverted compared to the previous lemmas. In this lemma, the adversary
assumes control first with a capability for the closure creator
\texttt{g1}. This means we need to check all arguments are safe
to use and that we clean up before returning in the end. The inversion
of control poses an interesting challenge when it comes to reasoning
about the adversaries local state during the execution of
\texttt{f4}. During the execution of \texttt{f4} and the callbacks, the
adversary should not be able to rely on its local state from before
the call of \texttt{f4}. This is easily done by revoking all the
temporary regions of the world given at the start of \texttt{f4}. The
real challenge comes when \texttt{f4} returns and we need to establish
that the protocols of the adversaries local state has not been
violated. This can be done by reinstating all of the adverseries old
regions, but we can only do this if the old adverserial regions are
not masked by permanent regions. As illustrated in
\figurename~\ref{fig:worlds-in-pf-g1}, we know this must be the
case. Whenever we are given a world, we get a public future world
where it is not allowed to use permanent regions for masking and in
the private future worlds we construct, we know it is not done
either. The return pointer \texttt{f4} gets from the adversary is
valid in future worlds of $W_1$, so in order to use it to argue it is
safe to return $W_6$ needs to be a public future worlds of $W_1$. We
can construct this as a private future world of $W_5$, but only
because we can reinstate the temporary regions of $W_1$ with the same
region names.

\begin{figure}[hptb]
  \centering
\begin{tikzpicture}[main node/.style={}, node distance=1.5cm]
  \node[main node] (1) {$W_1$};
  \node[main node] (label1) [below of=1,yshift=1cm] {\footnotesize{\textit{\texttt{f3} called}}};
  \node[main node,left] (given) [left of=1,yshift=0.05cm] {\parbox{1.5cm}{\footnotesize{Given:}}};

@@

  \node[main node] (6) [above right of=5]{$W_6$};
  \node[main node] (label6) [above of=6,yshift=-1cm] {\footnotesize{\textit{\texttt{f3} returns}}};
  \path(1) edge[draw=none] node [sloped, auto=false, allow upside down] {$\sqsubseteq^\priv$} (2)
       (2) edge[draw=none] node [sloped, auto=false, allow upside down] {$\sqsubseteq^\pub$} (3)
       (3) edge[draw=none] node [sloped, auto=false, allow upside down] {$\sqsubseteq^\priv$} (4)
       (4) edge[draw=none] node [sloped, auto=false, allow upside down] {$\sqsubseteq^\pub$} (5)
       (5) edge[draw=none] node [sloped, auto=false, allow upside down] {$\sqsubseteq^\priv$} (6);
\end{tikzpicture}
  \caption{Worlds in the proof of Lemma~\ref{lem:correctness-g1}}
  \label{fig:worlds-in-pf-g1}
\end{figure}
\begin{itemize}
\item Ticket dispenser
\item The awkward example and variants
\item A sandboxing example?

>>>>

In Lemma~\ref{lem:correctness-g1} the order of control has been
inverted compared to the previous lemmas. In this lemma, the adversary
assumes control first with a capability for the closure creator
\texttt{g1}. This means we need to check all arguments are safe
to use and that we clean up before returning in the end. The inversion
of control poses an interesting challenge when it comes to reasoning
about the adversaries local state during the execution of
\texttt{f4}. During the execution of \texttt{f4} and the callbacks, the
adversary should not be able to rely on its local state from before
the call of \texttt{f4}. This is easily done by revoking all the
temporary regions of the world given at the start of \texttt{f4}. When
\texttt{f4} returns, then the adversary is allowed to rely on its
local state again which means that we need to guarantee that the local
state is unchanged. To illustrate how this is done in the proof of
Lemma~\ref{lem:correctness-g1}, consider
\figurename~\ref{fig:worlds-in-pf-g1}. The return pointer given to
\texttt{f4} might be local and it is capability safe in $W_1$, so it
can only be used in public future worlds. Let us see how $W_6$ is
constructed as a public future world. The worlds $W_2$ and $W_4$ are constructed by revoking all
temporary regions and adding a $\iota^\pwl$ region for the stack we
pass away and static temporary regions (a region that accepts exactly
one memory) for our local stack as well as the adversaries local stack.
The given worlds $W_3$ and $W_5$ are public future worlds. None of
these worlds can have masked the temporary regions of $W_1$ with permanent
regions. $W_6$ is constructed by revoking the temporary regions of
$W_5$ and reinstating the temporary regions of $W_1$. This makes $W_6$
a public future world of $W_1$. It is therefore safe to use the return
pointer to return to the adversary.

\begin{figure}[t]
  \centering
\begin{tikzpicture}[main node/.style={}, node distance=1.5cm]
  \node[main node] (1) {$W_1$};
  \node[main node] (label1) [below of=1,yshift=1cm] {\footnotesize{\textit{\texttt{f3} called}}};
  \node[main node,left] (given) [left of=1,yshift=0.05cm] {\parbox{1.5cm}{\footnotesize{Given:}}};

@@

  \node[main node] (6) [above right of=5]{$W_6$};
  \node[main node] (label6) [above of=6,yshift=-1cm] {\footnotesize{\textit{\texttt{f3} returns}}};
  \path(1) edge[draw=none] node [sloped, auto=false, allow upside down] {$\sqsubseteq^\priv$} (2)
       (2) edge[draw=none] node [sloped, auto=false, allow upside down] {$\sqsubseteq^\pub$} (3)
       (3) edge[draw=none] node [sloped, auto=false, allow upside down] {$\sqsubseteq^\priv$} (4)
       (4) edge[draw=none] node [sloped, auto=false, allow upside down] {$\sqsubseteq^\pub$} (5)
       (5) edge[draw=none] node [sloped, auto=false, allow upside down] {$\sqsubseteq^\priv$} (6);
\end{tikzpicture}
  \caption{Illustration of some of the worlds in the proof of
    Lemma~\ref{lem:correctness-g1}. The first row of worlds are worlds
  we construct during the proof. The second row of worlds are worlds
  we are given during the proof. $W_6$ is constructed such that $W_6
  \futurewk W_1$.}
  \label{fig:worlds-in-pf-g1}
\end{figure}
\begin{itemize}
\item Ticket dispenser
\item The awkward example and variants
\item A sandboxing example?
